\chapter{Conclusion and Future Work}

\section{Summary}

This thesis studied learning-augmented algorithms for online bipartite
matching by first building a common experimental foundation and then
explaining what it revealed. On a single harness --- validated against
the published Borodin et al.~study (Chapter 3) --- we compared the
advice-free baselines, MinPredictedDegree and its augmentations, and the
test-and-fallback algorithms, across synthetic graphs, six real-world
graphs, and real request traces (Chapters 4--7). Across every setting
the same picture appeared: the advice-free baseline is already
near-optimal on average-case inputs, so the \emph{consistency} upside of
a good prediction is small; unguarded prediction-following crashes
\emph{below} the baseline under bad predictions; and the value of the
sophisticated algorithms is downside protection, delivered by either a
structural or an adaptive robustness mechanism with distinct trade-offs.
We sharpened the sub-questions --- that the (small) loss is governed by
a Kendall-\(\tau\) order error rather than magnitude (Chapter 5,
crediting the prior order-dependence result of ACI), and that the
adaptive test has a threshold-calibration pathology and a resolution
limit (Chapter 6) --- and we honestly reported the directions that did
\emph{not} pan out: learning the predictor for extra performance, and
finding a serving regime where predictions genuinely help (Chapter 8).

Finally, we argued the recurring wall is \emph{necessary} (Chapter 9):
no test-and-fallback algorithm whose test inspects only a sublinear
prefix can be both consistent and robust on strong-baseline matching,
because the structure that makes the prefix distribution-test feasible
is the structure that makes the baseline near-optimal. Experiments
discover the wall; theory explains why it must be there. The single
sentence that unifies the thesis is: \textbf{on average-case online
matching, predictions are robustness insurance rather than a performance
lever --- and where you can test the advice, you do not need it.}

\section{Limitations}

\begin{itemize}
\tightlist
\item
  \textbf{Input model.} We work in the known-i.i.d. model. Because
  Known-I.I.D. \(\le\) Random-Order in difficulty, the algorithms'
  guarantees carry over, but the empirical wall is an
  \emph{average-case} statement; we do not claim it for adversarial
  arrival order.
\item
  \textbf{Test model.} Following the original authors, the
  test-and-fallback experiments and the theorem's reduction use an
  empirical-\(\ell_1\) surrogate for the (unimplemented) distribution
  tester.
\item
  \textbf{Prediction-object heterogeneity.} The degree- and
  histogram-prediction families do not map onto every graph, which is
  why they are reported in parallel panels rather than one table.
\item
  \textbf{Data breadth.} Each real modality is exercised by one trace.
\item
  \textbf{Theory scope and status.} The impossibility theorem is
  strong-form in the \(r=\Theta(n)\) regime --- the regime with an
  upside to contest --- and one routine step of its proof (the
  witness-instance construction) plus a final review remain before it is
  fully typeset; a version biting at constant \(r\) is open.
\end{itemize}

\section{Future work}

The thesis brackets, rather than resolves, the settings in which
predictions might genuinely help online matching, and these are the
natural next directions.

\begin{itemize}
\tightlist
\item
  \textbf{Beyond average-case inputs.} The wall is an average-case
  phenomenon. Adversarial or non-stationary arrival orders, where the
  advice-free baseline is provably far from optimal, are where
  predictions should carry real value --- and where a \emph{positive}
  counterpart to this thesis's impossibility might be proved.
\item
  \textbf{Beyond throughput.} Objectives on which the baseline is not
  near-optimal --- tail latency, per-type fairness, migration or
  recompute cost --- may admit genuine with-predictions gains that the
  goodput objective forecloses; our serving SLO probe (Chapter 8) closed
  the simplest such attempt but not the space.
\item
  \textbf{Completing and strengthening the theory.} Closing the
  remaining construction step and the review; extending the
  impossibility to constant \(r\) or to other prediction objects; and a
  matching \emph{upper} bound --- for instance, showing that a
  recalibrated-threshold algorithm is optimal among test-and-fallback
  schemes --- would turn Chapter 9 into a tight characterization.
\item
  \textbf{Better tests, honestly.} Whether a super-linear or amortized
  test, or a test that reuses online decisions as samples, can escape
  the tolerant-testing barrier is an open and practically motivated
  question.
\end{itemize}

\section{Closing}

Predictions are a powerful tool for online algorithms, but this thesis
is a study of their \emph{limits} on one well-understood problem. On
average-case online matching, the honest verdict is that a cheap,
order-faithful predictor already captures nearly all there is to
capture, that the sophisticated machinery earns its keep as insurance
rather than as performance, and that no sublinear test can do better ---
because the very feasibility of the test is a signal that the prediction
was not needed. Recognizing where predictions cannot help is, we hope,
as useful as knowing where they can.
